%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.3 (9/9/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:1
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{llncs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{subfig}

\usepackage{hyperref} % For hyperlinks in the PDF
\usepackage{graphicx}

\newcommand{\raspberry}{\textsf{raspberry pi}\ }
\newcommand{\tablet}{\textsf{tablet}\ }
\newcommand{\iMac}{\textsf{iMac}\ }
\newcommand{\laptop}{\textsf{laptop}\ }
\newcommand{\blade}{\textsf{blade}\ }
\newcommand{\raspberrynsp}{\textsf{raspberry pi}}
\newcommand{\tabletnsp}{\textsf{tablet}}
\newcommand{\iMacnsp}{\textsf{iMac}}
\newcommand{\laptopnsp}{\textsf{laptop}}
\newcommand{\bladensp}{\textsf{blade}}

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{A Cross-Platform Assessment of Energy Consumption in Evolutionary Algorithms}
\subtitle{Towards Energy-Aware Bioinspired Algorithms}

\author{F. Fernández de Vega \inst{1} \and F. Ch\'avez\inst{1} \and J. D\'iaz\inst{1} \and J.A. Garc\'ia\inst{1} \and P.A. Castillo\inst{2} \and Juan J. Merelo\inst{2} \and C. Cotta\inst{3}\\
\institute{ Universidad de Extremadura \\
{\email{\{fcofdez,fchavez,mjdiaz,jangelgm\}@unex.es}} \\
\and ETSI Inform\'atica, Universidad de Granada\\
{\email{\{pacv,jmerelo\}@ugr.es}} \\ 
\and ETSI Inform\'atica, Campus de Teatinos, Universidad de M\'alaga\\
{\email{ccottap@lcc.uma.es}} \\
}}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Insert title

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\begin{abstract}
Energy consumption is a matter of paramount importance in nowadays
environmentally conscious society. It is also bound to be a crucial
issue in light of the emergent computational environments arising from
the pervasive use of networked handheld devices and
wearables. Evolutionary algorithms (EAs) are ideally suited for this
kind of environments due to their intrinsic flexibility and
adaptiveness, provided they operate on viable energy terms.  In this
work we analyze the energy requirements of EAs, and particularly one
of their main flavours, genetic programming (GP), on several
computational platforms and study the impact that parametrisation has
on these requirements, paving the way for a future generation of
energy-aware EAs.  As experimentally demonstrated, handheld devices
and tiny computer models mainly use for educational purposes may be
the most energy efficient ones when looking for solutions by means of
EAs.
\end{abstract}

\noindent \textbf{Keywords:} Green computing, energy-aware computing,
performance measurements, evolutionary algorithms. 

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\input{introduction}

\input{literature}

\input{methodology}

\input{results}

\input{conclusions}

\subsubsection*{Acknowledgements}
\sloppypar We acknowledge support from 
Spanish Ministry of Economy and Competitiveness and European Regional
Development Fund (FEDER) under project EphemeCH
(TIN2014-56494-C4-\{1,2,3\}-P),  
from University of Granada, PROY-PP2015-06 (Plan Propio 2015 UGR), 
from Junta de Andaluc\'{\i}a under project DNEMESIS (P10-TIC-6083),
from Universidad de M\'alaga, Campus de Excelencia Internacional
Andaluc\'{\i}a Tech,  
from Junta de Extremadura FEDER, project GR15068 and FP7-PEOPLE-2013 IRSES Grant 612689 ACoBSEC.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
 
\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

%\bibitem{smith}{Smith, S. F. (1980). \textit{A Learning System Based on Genetic Adaptive Algorithms}. PhD thesis, Computer Science Department, University of Pittsburgh.}
%\bibitem{soule}{Soule, T. and Foster, J. A. (1998a). Effects of code growth and parsimony pressure on populations in genetic programming. \textit{Evolutionary Computation}, 6(4):293-309.}
%\bibitem{tackett}{Tackett, W. A. (1994). \textit{Recombination, Selection, and the Genetic Construction of Computer Programs}. PhD thesis, University of Southern California, Department of Electrical Engineering Systems.}
%\bibitem{kalganova}{Kalganova, T. and Miller, J. (1999). Evolving more efficient digital circuits by allowing circuit layout evolution and multi-objective fitness. In Stoica, A., Keymeulen, D., and Lohn, J., editors, \textit{Proceedings of the 1st NASA/DoD Workshop on Evolvable Hardware (EH’99)}, pages 54-63, Piscataway, NJ. IEEE.}
%bibitem{zhang}{Zhang, B.-T. and Muhlenbein, H. (1995). Balancing accuracy and parsimony in genetic programming. \textit{Evolutionary Computation}, 3(1):17-38.}
%\bibitem{nedja}{Nedja, N., Abraham, A., and de Macedo Mourelle, L. editors (2006). \textit{Genetic Systems Programming}. Springer.}
%\bibitem{silva}{Silva, S. and Almeida, J. (2003). Dynamic maximum tree depth. In \textit{Genetic and Evolutionary Computation, GECCO-2003}, pages 1776-1787, Chicago. Springer.}
\bibitem{green-computing}{Hooper, A. (2008). Green
    computing. Communication of the ACM, 51(10), 11-13.}
\bibitem{energyaware}{Barroso, L. A., Hölzle, U. (2007). The case for energy-proportional computing. Computer, (12), 33-37.}
\bibitem{HVAC}{Fong, K. F., Hanby, V. I., Chow, T. T. (2006). HVAC system optimization for energy management by evolutionary programming. Energy and Buildings, 38(3), 220-231. Elsevier}
\bibitem{chiller}{Lee, W. S., Chen, Y. T., Kao, Y. (2011). Optimal chiller loading by differential evolution algorithm for reducing energy consumption. Energy and Buildings, 43(2), 599-604. Elsevier}
\bibitem{dispatch}{Fadaee, M., Radzi, M. A. M. (2012). Multi-objective optimization of a stand-alone hybrid renewable energy system by using evolutionary algorithms: A review. Renewable and Sustainable Energy Reviews, 16(5), 3364-3369.}
\bibitem{self}{Cotta, C. Fernández, A. J., Fernández de Vega, F., Chávez, F., Merelo, J.J.,Castillo, P.A., Camacho, D., Bello, G:
Ephemeral Computing and Bioinspired Optimization. IJCCI (ECTA) 2015: 319-324}
\bibitem{energy-aware}{Yoo, C. M. K. S. (2011). Energy-Aware System Design. Springer.}
\bibitem{cache}{Díaz, J., Risco, J.L., Colmenar, J.M., Multi-objective optimization of energy consumption and execution time in a single level cache memory for embedded systems. Journal of Systems and Software, vol 111, pp 200-212.  Elsevier}
\bibitem{scaling}{Bansal, N., Kimbrel, T., Pruhs, K. (2004, October). Dynamic speed scaling to manage energy and temperature. In Foundations of Computer Science, 2004. Proceedings. 45th Annual IEEE Symposium on (pp. 520-529). IEEE.}
\bibitem{dynamic-scaling}{Albers, S. (2011). Algorithms for dynamic speed scaling. In LIPIcs-Leibniz International Proceedings in Informatics (Vol. 9). Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik.}
\bibitem{energy-efficient}{Albers, S. (2010). Energy-efficient algorithms. Communications of the ACM, 53(5), 86-96.}
\bibitem{michalewicz}{Zbigniew Michalewicz,(1996). Genetic Algorithms + Data Structures = Evolution Programs. Third Edition, Springer.}
\bibitem{cantupaz}{Cantu-Paz, E. (2000). Efficient and Accurate Parallel Genetic Algorithms, Kluwer Academic Publishers.}
\bibitem{fcofdez}{Fernández, F., Tomassini, M., Vanneschi, L. (2003). An empirical study of multipopulation genetic programming. Genetic Programming and Evolvable Machines, 4(1), 21-51. Springer.}
%\bibitem{web_algoritmo}{Dennis Cormier and Sita Raghavan. http://people.sc.fsu.edu/~jburkardt/cpp\_src/ simple\_ga/simple\_ga.html.}
%\bibitem{complexity}{Papadimitriou, C. H. (2003). Computational complexity (pp. 260-265). John Wiley and Sons Ltd..}
\bibitem{ephemeral}{Cotta, C., Fernández, A., Fernández de Vega, F, Chávez, F., Merelo, J.J., Castillo, P., Camacho, D., Bello, G., Ephemeral computing and bioinspired optimization:  challenges and opportunities.  Proceedings ECTA 2015. pp. 319-324.}
\bibitem{paba}{Fernández de Vega, F., Pérez, J. I. H. (2012). Parallel Architectures and Bioinspired Algorithms. Springer.}
%\bibitem{powertutor}{Gordon, M., Zhang, L., Tiwana, B., Dick, R., Mao, Z. M., Yang, L. (2013). PowerTutor: a power monitor for android-based mobile platforms.}
\bibitem{powertutor2}{Zhang, L., Tiwana, B., Qian, Z., Wang, Z., Dick, R. P., Mao, Z. M., Yang, L. (2010, October). Accurate online power estimation and automatic battery behavior based power model generation for smartphones. In Proceedings of the eighth IEEE/ACM/IFIP international conference on Hardware/software codesign and system synthesis (pp. 105-114). ACM.}
%\bibitem{lilgp}{Punch, B., Zongker, D. (1998). Lil-gp genetic programming system. Michigan State University GARAGe.}
\bibitem{barroso2007case}{Barroso, L.A. and H{\"o}lzle, U. (2007). The case for energy-proportional computing. Computer, Vol. 12, pp. 33-37. IEEE}
\bibitem{koza}{Koza, J. R. (1992). Genetic programming: on the programming of computers by means of natural selection (Vol. 1). MIT press.}
\bibitem{pool}{García-Valdez, M., Trujillo, L., Merelo, J. J., de Vega, F. F., Olague, G. (2015). The EvoSpace model for pool-based evolutionary algorithms. Journal of Grid Computing, 13(3), 329-349.}
\end{thebibliography}


%----------------------------------------------------------------------------------------

\end{document}
